### 2018-03-23 Writing/draft.tex
changed CD equation
Added the fact that our simulation protocol is not biological

\subsubsection{Character Difference ($CD$)}
\begin{equation}
    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
\end{equation}

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each characters states for the each $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
%TG: Justification for the standardisation: basically, it makes comparisons faster and Fitch-like.
This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
It also greatly improved the speed of our algorithm implementation to compare the characters.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
%MB: I don't get this part. What do you mean by standardized? And why are you doing this? Are you just editing the symbols?
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).

@@

% Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
% %MB: Do we? I'm not sure that I would. In practice, this would probably force morphologists to remove most of their characters.
% %MB: Making these kinds of recommendations leads to the risk that all you do is introduce biases into the dataset in another way.
% %TG: Fair enough, I'm happy to drop this whole paragraph.
% However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
% %MB: How small? Be specific.
% Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e. as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).

@@

% This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
%MB: This is serious hedging ("we're not going to... but we're going to"). Could probably skip the above line.
%TG: fair point
% The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
% When many characters are homoplasic (high character difference) parsimony will deal with this information by generating longer trees (i.e. ) (which will be more likely to be discarded in the tree search).
%MB: This doesn't make sense. Longer than one? How is generating longer trees coping? What do you mean by "coping"? 
% Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
% These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
%MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
%TG: I've rephrased the parargaph below. This is the idea (a bit arms waving, I agree): for a given topology, when there is high homoplasy, parsimony will give the tree a high score (may changes per branches - even though branch length is not a parsimony parameter) and probably discard the topology. For the same topologyn and high homoplasy, Bayesian will stretch the branches (accounting for the many changes per branches) which will decrease the likelihood score but less fast than MP (because branch length is only *one* of the parameters accounting for the likelihood score). Also, in Bayesian, an decrease in likelihood score does not necessary mean that the tree will be discarded.
The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.

@@

% Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis. TG: moved to intro.
%MB: From a structural point of view, this line is out of place even. It makes a general point, but deep in the subsection. Such statements are better made earlier so a reader knows what the point of the section is.
It is important to note that the two other sources of character correlation could also be present in our simulations although they were not explicitly modelled:
(1) evolutionary correlation is implied by simulating the characters based on Birth-Death trees;
and (2) intra-organismal correlation could also be present in the matrices for those characters randomly simulated but sharing similar evolutionary simulation regimes (i.e. creating ``modules'' of characters).
However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.

\subsection{Limitations}
First, simulating evolutionary history is complex.
Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible.
This can lead to spuriously generated discrete morphological matrices
However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.

Second, measuring and modifying character correlation is difficult. 
Increasing overall characters difference so that each pairwise differences is equal to 1 becomes more and more impossible with a large number of characters.
In our algorithm, we maximised characters difference by replacing a characters $A$ highly correlated to another other character $B$ by a randomly chosen character $X$.
This reduces the correlation between $A$ and $B$ but increases the correlation between $A$ and $X$!
Additionally, when minimising $n$ character differences, the matrix will end containing only one character duplicated $n$ times (which will generate a unresolved tree by definition).
Therefore we used an algorithm that only ``maximises'' or ``minimises'' character differences on average (see Fig. \ref{Fig:modif_matrix}).
Although this was sufficient to generate detectable differences in topologies, it is not excluded that some randomly ``maximised'' or ``minimised'' matrices were actually not different than the ``randomised'' ones.

@@

@@

However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
%MB: But at what point? How much correlation is needed to affect it? TG: added values.
%TG: do we want to suggest something here to close the conclusion? I agree we don't wanna be presecriptive Bristolians but it'll be nice to finsh with something like: "we're not saying it's a problem per se but maybe check your data?"

>>>>

\subsubsection{Character Difference ($CD$)}
\begin{equation}
    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
\end{equation}

%MB: Why divide by 1/2 when you can just multiply by 2?
%TG: fair point. I've fixed it in the demonstration as well (supplementaries)

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
%TG: Justification for the standardisation: basically, it makes comparisons faster and Fitch-like.
This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
It also greatly improved the speed of our algorithm implementation to compare the characters.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

@@

% However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
% Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e. as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
%MB: Agreed that that is noteworthy. I suspect I know why this happens (and you might have explained it earlier). It might actually be a statistical problem with our approach. When character similaritiy is maximized, how well resolved are the trees? 
%TG: I've moved that to a "potential application" paragraph in the discussion (please change the title of it, I had zero inspiration and find it ugly as feck!)

@@

%MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
%TG: I've rephrased the parargaph below. This is the idea (a bit arms waving, I agree): for a given topology, when there is high homoplasy, parsimony will give the tree a high score (may changes per branches - even though branch length is not a parsimony parameter) and probably discard the topology. For the same topologyn and high homoplasy, Bayesian will stretch the branches (accounting for the many changes per branches) which will decrease the likelihood score but less fast than MP (because branch length is only *one* of the parameters accounting for the likelihood score). Also, in Bayesian, an decrease in likelihood score does not necessary mean that the tree will be discarded.
The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.

@@

It is important to note that the two other sources of character correlation could also be present in our simulations although they were not explicitly modelled:
(1) evolutionary correlation is implied by simulating the characters based on Birth-Death trees;
and (2) intra-organismal correlation could also be present in the matrices for those characters randomly simulated but sharing similar evolutionary simulation regimes (i.e. creating ``modules'' of characters).
However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.

\subsection{Limitations}
%MB: I think what this section _really_ needs is an assessment of the method used to simulate correlations. It was basically accomplished by duplicating characters. It wasn't accomplished by evolving some characters with a particular covariance model (which probably would've been better). We kind of faked-up that model. 
%TG: agree.

TODO

First, simulating evolutionary history is complex.
Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible.
This can lead to spuriously generated discrete morphological matrices
However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.

%TG: insisting on the shittiness of generatin correlation
Second, measuring and modifying character correlation is difficult. 
In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
%MB: We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
%MB: We didn't look at ranking and we didn't look at competing correlations. There's a lot more that could be done here so, we might talk about that.
%TG: not sure what you mean by the ranking stuff but it could definitely fit here.
However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e. some sets of characters can be highly correlated but effectively describe independent information).
Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.

@@

\subsection{Potential applications}
%TG: explaining that our "scenarios" are biased bootstraps
Effectively, our simulation protocol bootstraps our data ``with bias''.
In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.

%TG: explaining that we could use it for "clever" bootstrapping in big matrices
It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling (i.e. the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.

@@

However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
%MB: But at what point? How much correlation is needed to affect it? TG: added values.

%MB: Part of it is: we expect it to be a problem, but _when_ do we expect it to be a problem. There's the issue of "check your data", but it only raises the question: how? I think this is where we suggest avenues for future research (either here or move it up to the discussion).
